systemfit: A Package to Estimate\\ Simultaneous Equation Systems in \proglang{R}

### 2005-11-08 paper/trunk/systemfit.tex
made seemingly massive changes... but hopefully not for the worse...
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@586 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@

%% an abstract and keywords
\Abstract{Many statistical analysis are based on models that consist
  of  more than one equation. If these models contain relationships
  among the various equations, ignoring these relationships can result
  in biased and possibly inconsistent parameter estimates. If
  relationships exist among the equations, the sets of equations
  should be fitted simultaneously rather than equation by equation.
  The package \pkg{systemfit} provides the capability to estimate
  systems of linear and non-linear equations, using limited and
  full-information methods within the \proglang{R} programming
  environment.  }

\Keywords{R, simultaneous equation system, seemingly unrelated regression,
   two-stage least squares, three-stage least squares}

@@

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{Estimating nonlinear equation systems}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% The \pkg{systemfit} package also contains a method for fitting
% nonlinear systems of equations.

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsection{Nonlinear parameter estimation}
% \label{sec:nonlinear_estimation}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% A system of nonlinear equations can be written as:

% \begin{equation}
%   \label{eq:non_linear_eq_1}
%   \epsilon_{t} = q( y_t, x_t, \beta )
% \end{equation}
% %%% It is not clear what these variable are. 
% %%% Adding an equation representing only a single equation
% %%% would make things clearer.

% \noindent and

% \begin{equation}
%   \label{eq:non_linear_eq_2}
%   z_{t} = Z( x_t )
% \end{equation}

% where $\epsilon_{t}$ are the residuals from the y observations and
% $z_{t}$ are the function evaluated at the parameter estimates.

% \textbf{you really need to check this!!!}

% The objective functions for the methods are:

% \begin{center}
% \begin{tabular}{|l|c|c|c|} \hline
%   Method & Instruments & Objective Function & Covariance of $\theta$ \\ \hline  
%   OLS & No & $r'r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline
%   SUR & No & $r'(diag(S)_{OLS}^{-1}\otimes I)r$ & $(X(S^{-1}\otimes I)X)^{-1}$ \\ \hline
%   2SLS & Yes & $r'(I \otimes W)r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline 
%   3SLS & Yes & $r'(S_{2SLS}^{-1} \otimes W)r$ & $(X(diag(S)^{-1}\otimes W)X)^{-1}$ \\ \hline
% \end{tabular}
% \end{center}

% where, $r$ is a column vector for the residuals for each equation,
% %% what is the difference to \epsilon?
% $X$ is matrix of the partial derivatives of the dependent variable 
% with respect to the parameters $\left( \frac{ \partial y }{ \theta} \right)$,
% %% is this correct?
% %% in the linear section X is used for the regressors, 
% %% thus we should use a different variable name here.
% $W$ is a matrix of the instrument variables, $Z(Z'Z)^{-1}Z$, $Z$ is a
% matrix of the instrument variables, and $I$ is an $n \times n $
% identity matrix and $S$ is the estimated variance-covariance matrix between the
% equations

% \begin{equation}
%   \label{eq:non-linear_varcov}
%   \hat{\sigma}_{ij} = (\hat{e}_i' \hat{e}_j) / \sqrt{(T - k_i)*(T- k_j)} 
% \end{equation}

% \textbf{The residual variance-covariance matrix can be calculated in 
% different ways (section~\ref{sec:residcov}). 
% It should be relatively easy to implement this also in nlsystemfit(), 
% e.g. the function 'calcRCov' in systemfit() could be moved outside
% systemfit() that it can be used also by nlsystemfit().}

% \textbf{You need to clear this up.}

% % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % \subsection{Using nlsystemfit}\label{sec:UsingnlSystemfit}
% % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% % \code {nlsystemfit} fits a set of structural nonlinear equations using
% % Ordinary Least Squares (OLS), Seemingly Unrelated Regression (SUR),
% % Two-Stage Least Squares (2SLS), Three-Stage Least Squares (3SLS) using
% % the objective functions described in section
% % \ref{sec:nonlinear_estimation}.

% % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % \subsubsection{Standard usage}

% % Similar to calling the \code{systemfit} function, \code{nlsystemfit}
% % is called with a minimum of three arguments,

% % \code{
% % R> nlsystemfit( method, eqns, start )
% % }

% % where \code{method} is one of the following estimation methods: "OLS",
% % "SUR", "2SLS", or "3SLS", \code{eqns} is a list of equations similar
% % to those described for \code{systemfit}, and \code{start} is a list of
% % starting values of the parameter estimates.

% % % The mandatory argument \code{eqns} is a list of the equations  to
% % % estimate.  Each equation is a standard formula in \proglang{R}.  It
% % % starts with a dependent variable on the left hand side.  After a tilde
% % % ($~$) the regressors are listed  \footnote{For Details see the
% % %   \proglang{R} help files to \code{formula}}.

% % This is now demonstrated using an example: \\

% % % \code{
% % % R> library( systemfit ) \\
% % % R> data( ppine ) \\
% % % R> hg.formula <- hg $\sim$ exp( h0 + h1*log(tht) + h2*tht$\hat$2 + h3*elev + h4*cr) \\
% % % R> dg.formula <- dg $\sim$ exp( d0 + d1*log(dbh) + d2*hg + d3*cr + d4*ba  ) \\
% % % R> labels <- list( "height.growth", "diameter.growth" )\\
% % % R> inst <- $\sim$ tht + dbh + elev + cr + ba\\
% % % R> start.values <- c(h0=-0.5, h1=0.5, h2=-0.001, h3=0.0001, h4=0.08,\\
% % % R+   d0=-0.5, d1=0.009, d2=0.25, d3=0.005, d4=-0.02 )\\
% % % R> model <- list( hg.formula, dg.formula )\\
% % % R> model.3sls <- nlsystemfit( "3SLS", model, start.values, data=ppine,\\
% % % R+   eqnlabels=labels, inst=inst ) \\
% % % R> summary( model.3sls ) \\
% % % }

% % The nlsystemfit function relies on \code{nlm} to perform the
% % minimization of the objective functions and the \code{qr} set of
% % functions. If the user does not have a set of estimates for the
% % initial parameters, it is suggested using linearized forms in one of
% % the linear methods, or simply using \code{nls} to obtain estimates for
% % the equations.

% % The outputs are similar to those for a systemfit object.

% \section{CVS Revision Log}
% \begin{verbatim}
% $Log$
% Revision 1.18  2005/11/07 14:08:56  henningsena
% a few minor changes (predominantly adding a "\ " after a period (.) that does not indicate an end of a sentence)
%
% Revision 1.17  2005/11/07 10:21:59  henningsena
% just changed some line breaks (deleted new ones and restored deleted ones) to make it easier to find the changes that Jeff committed in the last revision(s)
%
% Revision 1.16  2005/11/02 18:31:45  hamannj
% small last changes
%
% Revision 1.15  2005/11/02 18:15:40  hamannj
% small changes

>>>>

\usepackage{lineno}

\linenumbers

@@

%% an abstract and keywords
\Abstract{

  Many statistical analyses are based on models containing systems of
  structurally related equations. If the cross-equation disturbances
  are correlated, full information methods are required
  \citep{zellner62}. If exogenous variables are stochastically
  dependent on the disturbances in the system
  \citep{zellner62b,theil71}, limited information methods should be
  used for parameter estimation. The package \pkg{systemfit} provides
  the capability to estimate systems of linear and non-linear
  structural equations for limited (ordinary least squares (OLS), two
  stage-least squares (2SLS)) and full-information
  (seemingly-unrelated regression (SUR) and three-stage least squares
  (3SLS)) methods within the \proglang{R} programming environment.}

\Keywords{R, simultaneous equations estimation, ordinary least
  squares, seemingly unrelated regression, two-stage least squares,
  three-stage least squares}

@@

% \section{CVS Revision Log}
% \begin{verbatim}
% $Log$
% Revision 1.19  2005/11/08 06:31:48  hamannj
% made seemingly massive changes... but hopefully not for the worse...
%
% Revision 1.18  2005/11/07 14:08:56  henningsena
% a few minor changes (predominantly adding a "\ " after a period (.) that does not indicate an end of a sentence)
%
% Revision 1.17  2005/11/07 10:21:59  henningsena
% just changed some line breaks (deleted new ones and restored deleted ones) to make it easier to find the changes that Jeff committed in the last revision(s)
%
% Revision 1.16  2005/11/02 18:31:45  hamannj
% small last changes
%
% Revision 1.15  2005/11/02 18:15:40  hamannj
% small changes

### 2005-11-08 paper/trunk/systemfit_intro.tex
made seemingly massive changes... but hopefully not for the worse...
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@586 591db2bd-e0a0-4d22-8c19-2192dbb7da31

Many theoretical models consist of more than one equation. If these
models are econometrically estimated, it may not be appropriate to
estimate the equations individually, but all together, because the
disturbance terms of these equations are likely to be contemporaneously
correlated. %\citep{theil71} (make sure this is the correct paper).
Estimating all equations simultaneously, taking the
covariance structure of the residuals into account, leads to efficient
estimates.
Ignoring this correlation leads to an inefficient
parameter estimation \citep{zellner62}.

Another reason to estimate an equation system simultaneously are
cross-equation parameter restrictions. These restrictions can be
tested and/or imposed only in a simultaneous estimation
approach especially the economic theory suggests many cross-equation
restrictions.

The \pkg{systemfit} package provides the capability to estimate
linear and non-linear equation systems in \proglang{R}.
Although linear and non-linear equation systems can be estimated
with several other statistical and econometric software packages
(e.g.\ \proglang{SAS}, \proglang{EViews}, \proglang{TSP}),
\pkg{systemfit} has several advantages.
First, all estimation procedures are publicly available in the source code.
Second, the estimation algorithms can be easily modified to meet specific
requirements.
Third, the (advanced) user can control many important details of the
estimation procedure that are generally not available in other software
packages.
On the other hand, these user options have reasonable defaults that also
beginners can easily use the provided functions.

>>>>

Many theoretical models consist of more than one equation. These
models can contain variables that appear on the left-hand side in one
equation and on the right-hand side of another equation
(simultaneity-bias). The system can have non-zero off-diagonal
covariance elements resulting from related disturbnaces
(contemporaneous correlation). Ignoring contemporaneous correlation
leads to inefficient parameter estimates \citep{zellner62} and
ignoring simultaneity-bias can lead to inconsistent parameter
estimates.

% If these models are econometrically estimated, it may not be
% appropriate to estimate the equations individually, but all together,
% because the disturbance terms of these equations are likely to be
% contemporaneously correlated.  Estimating all equations
% simultaneously, taking the covariance structure of the residuals into
% account, leads to efficient estimates.  Ignoring this correlation
% leads to an inefficient parameter estimation \citep{zellner62}.

Another reason to estimate an equation system simultaneously are
cross-equation parameter restrictions. These restrictions can be
tested and/or imposed only in a simultaneous estimation approach
especially the economic theory suggests many cross-equation
restrictions. (But we don't say why this is important...) Are we
imposing restrictions in the parameters, or the variance-covariance
matrix or both and for what pupose?

For all of the methods developed in the package, the disturbnaces of
the individual equations are assumed to be independent and identially
distributed (iid).  
In the future, we would like to add the ability to fit equations were
the disturbances are serially correlated (wikins 1969).

The \pkg{systemfit} package provides the capability to estimate
linear and non-linear equation systems in \proglang{R}.
Although linear and non-linear equation systems can be estimated
with several other statistical and econometric software packages
(e.g.\ \proglang{SAS}, \proglang{EViews}, \proglang{TSP}),
\pkg{systemfit} has several advantages.
First, all estimation procedures are publicly available in the source code.
Second, the estimation algorithms can be easily modified to meet specific
requirements.
Third, the (advanced) user can control estimation details generally
not available in other software packages by overriding reasonable defaults.
